### 2018-11-13 Sabanci_talk/Sabanci_talk.tex
Finalish version

\documentclass[xcolor=dvipsnames,t]{beamer}
%\documentclass[xcolor=dvipsnames,handout,t]{beamer}
% use handout in document class to stop transitions

\mode<presentation>
{
  \usetheme{Madrid}%{Madrid}
  % or ..AnnArbor, Boadilla, CambridgeUS, Copenhagen, default, Frankfurt, Goettingen, Hannover
  %       Montpellier, PaloAlto, Rochester, Szeged,

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}
\usepackage{color,empheq}
\usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[english]{babel}
\usepackage{xmpmulti,animate}
\usecolortheme{seahorse} %this makes the colors less strong.

\usepackage{times}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.
\usepackage{graphicx}
%\usepackage{tcolorbox}
\usepackage{tikz,lipsum,lmodern}
\usepackage[most]{tcolorbox}
\usepackage{amsfonts,amssymb,amsmath,mathtools}
      % use Times fonts if available on your TeX system
\usepackage{epsfig}
\usepackage{hyperref}
\usepackage{changepage}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%\usepackage{animate,movie15,media9}
%\usepackage{multimedia}

\newcommand{\todo}[1]{\textcolor{orange}{\texttt{TODO: #1}}} 
\newcommand{\red}[1]{\textcolor{red}{#1}} 
\newcommand{\bl}[1]{\textcolor{blue}{#1}} 
\newcommand{\w}[1]{\textcolor{white}{#1}}

%\setlength{\droptitle}{10em}

\usetikzlibrary{shapes}

\title[] % (optional, use only with long paper titles)

%\subtitle
%{Presentation Subtitle} % (optional)

\title[Forecasting GRBs with GWs] % (optional, use only with long paper titles)
{\vspace{0mm}Forecasting Gamma-ray Bursts with Gravitational Waves}

\subtitle{} % (optional)
%\vspace{2cm}
\author[Sarp Ak\c{c}ay]{\vspace{-2cm}Sarp Ak\c cay\inst{1}\inst{2} }
\institute[FSU Jena - UCD] % (optional, but mostly needed)
{\vspace{-2.25cm}
  \inst{1}%
  FSU Jena %and University College Dublin
  \inst{2}%
  University College Dublin
  }

\date[Sabanc\i~Universitesi]{\vspace{-2.5cm}14 November 2018}
%\vspace{-3cm}
%\titlegraphic{
%  \vspace{1.75cm}  \includegraphics[width=1.0cm]{figs/ucd.png}%\hspace*{1cm}~\includegraphics[height=1cm]{figs/missing_inch.jpg} \hspace*{1cm}  \includegraphics[width=1.0cm]{figs/ucd.png}
%}

\subject{Talks}

\usefonttheme[onlymath]{serif}
\setbeamerfont{frametitle}{size=\huge}
\setbeamercolor{frametitle}{fg=Black,bg=White}

\newcommand{\Mag}{\textcolor{magenta}}
\newcommand{\Red}{\textcolor{red}}
\newcommand{\Blue}{\textcolor{blue}}
%\renewcommand{\c}{\cos}  \c{c} gives cecedille
\renewcommand{\t}{\theta}
\renewcommand{\b}{\bar}
\newcommand{\f}{\frac}
\newcommand{\bt}{\beta}
%\newcommand{\s}{\sin}
\newcommand{\ph}{\phi}
\newcommand{\g}{\gamma}
\newcommand{\nn}{\nonumber}
\newcommand{\la}{\lambda}
\newcommand{\al}{\alpha}
\newcommand{\La}{\Lambda}
\newcommand{\el}{\ell}
\newcommand{\h}{\hat}
\newcommand{\mrm}{\mathrm}
\newcommand{\ord}{\mathcal{O}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bef}{\begin{frame}}
\newcommand{\ef}{\end{frame}}
%\newcommand{\h}{\bar{h}}
\newcommand{\bs}{\begin{small}}
\newcommand{\es}{\end{small}}
\newcommand{\parallelsum}{\mathbin{\!/\mkern-5mu/\!}}
\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}
\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut for vertical spacing in tables
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
%------------------ for coloured boxes in math modes --------------------
% Syntax: \colorboxed[<color model>]{<color specification>}{<math formula>}
\newcommand*{\colorboxed}{}
\def\colorboxed#1#{%
  \colorboxedAux{#1}%
}
\newcommand*{\colorboxedAux}[3]{%
  % #1: optional argument for color model
  % #2: color specification
  % #3: formula
  \begingroup
    \colorlet{cb@saved}{.}%
    \color#1{#2}%
    \boxed{%
      \color{cb@saved}%
      #3%
    }%
  \endgroup
}
%%-------------------------------------------------------------------------

% tikz boxes
\newcommand\TBox[3][]{%
  \tikz\node[draw,ultra thick,red,text width=#2,align=left,#1] {#3};}

\setbeamertemplate{frametitle}[default][center]

\begin{document}
{
  \usebackgroundtemplate{\includegraphics[width=1\paperwidth,height=\paperheight]{figs/kilonova.jpg}}
  % image credit [University of Warwick/Mark Garlick]
  \begin{frame}
  \vspace{-2cm}
    \titlepage
  \end{frame}
} % these brackets are needed to insure that the background image is used ONLY in the title slide

\begin{frame}{Outline}
\begin{itemize}
 \item Motivation 
 \item \bl{Gamma-ray bursts}
 \item Neutron stars
 \item \bl{Gravitational waves}
 \item Compact binary \red{inspirals}
 \item Gravitational-wave detectors
 \item Gravitational-wave astronomy in the 2020s and the 2030s
 \item \red{Forecasting} prospects, event rates etc.
\end{itemize}

\end{frame}

\begin{frame}{Motivation: Multi-messenger Astronomy}
  \begin{itemize}
    \onslide<1->
    \item \red{\emph{Multi}}: two fundamentally different waves from common events.
    \item[]\quad Gravitational waves ({\bf GW}s): ``hearing''  the universe.
    \item[]\quad Electromagnetic waves (EWs): ``seeing'' the universe.
    \onslide<2->\item \red{GW170817}-\bl{GRB170817A}-\textcolor{brown}{AT2017gfo}: the ``event'' \\
    \quad inspiral + merger of two neutron stars (GWs) \\
    \ \ \ = short-hard gamma-ray burst + kilonova (EWs)
    \item[] Broad-band in gravitational waves: 20Hz - 2000Hz
    \item[] Ultra broad-band in electromagnetic waves: gamma rays to radio.
    \onslide<3->\item A simple question: \\
    {\small ``What can we offer to the astronomy community with future GW detections?''}
    \item[] \red{GRB/kilonova Forecasts!}
    \item[] Observe NOT 11 hours \underline{after} the merger, but ``live''
  \end{itemize}
\end{frame}

\begin{frame}{Gamma-ray bursts}{Most energetic explosions in the universe}
\begin{small}
    \begin{itemize}
      \item Discovered by Vela satellites in \red{1967}
      \onslide<2->\item Non-thermal spectrum: $10-10^4\,$\red{keV}
      \item Highly energetic $10^{48}-10^{\red{52}}\,$ergs 
      \onslide<3->{\item \bl{1000s} detected, $\sim 1$ per day}
      \onslide<4->\item Far away: $z\approx 1 \sim 8\,$billion years! %$9.4\sim 13\,$billion light years!
      \onslide<5->\item Afterglow from X-rays to radio \\
      over timescales of hours to months
      \onslide<6->\item \bl{Two types}: \red{short} and long
      \onslide<7->{\item Short GRBs:\\ 
      mergers of binary neutron stars (1989)\\
      \red{GW170817}-\bl{GRB170817A}}
    \end{itemize}
  \end{small}
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<2>\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.9cm,1.4cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_spectrum.jpg}};    
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.5cm,1.cm)}]current page.center) { Mon.Not.Roy.Astron.Soc. 433 (2013) 2739};
      \onslide<3>\node (img1 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.8cm,0.9cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/Fermi_GRB_sky2.png}};  
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(5.5cm,1.9cm)}]current page.center) {NASA/Fermi};
      \onslide<4>\node (img2 )[anchor=center,scale=1,opacity=1] at ([shift={(3.6cm,-0.15cm)}]current page.center) {\includegraphics[height=3.8cm]{figs/GRB_vs_redshift.png}};  % from:1401.5069
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.cm,1.4cm)}]current page.center) {Astrophys.J.Suppl. 211 (2014) 12};
      \onslide<5>\node (img3 )[anchor=center,scale=1,opacity=1] at ([shift={(3.4cm,-1.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_afterglow.jpg}}; 
      \onslide<6->\node (img4 )[anchor=center,scale=1,opacity=1] at ([shift={(3.2cm,-2.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_burst_durations.jpg}};
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(2.cm,-1.2cm)}]current page.center) {NASA/GSFC};
    \end{tikzpicture}
\end{frame}

\begin{frame}{Neutron Stars}{Most extreme stars in the universe}  
  \begin{small}
    \begin{itemize}
      \item Core-collapse supernova remnants
      \item Mass: $\sim $ \red{1} to \red{3} solar masses ($M_\odot$)
      \item Very \red{compact}: radius $\sim 9$ to 15\,km
      \onslide<2->\item High spins: up to 700 times/second
      \onslide<3->\item Extremely \bl{dense}: teaspoon $\sim 10^9\,$tons
      \onslide<3->{\item Large magnetic fields: $10^4$ to $10^{\red{11}}\,$T}
      \onslide<4->\item Hot: surface temperature $10^5 - 10^6\,$K
      \onslide<5->\item They can come in \red{pairs} (binary) \\
      \onslide<6->Orbit \red{decays} over time due\\
      $\Longrightarrow$ \underline{Gravitational wave emission}!\\
      Observations \red{match} theoretical \\predictions very well
      %Equation of state \\
      %Piecewise polytropes
      %\[ p = K_i \rho^{\Gamma_i},\quad (\rho_{i-1}\le \rho \le \rho_i) \hspace{8cm}\]
    \end{itemize}
  \end{small}
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<1>
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.2cm,1.5cm)}]current page.center) {\includegraphics[height=3.2cm]{figs/Core_collapseSXS.jpg}};    
	\node[anchor=center,scale=1,opacity=1] at ([shift={(5.4cm,2.9cm)}]current page.center) {\textcolor{white}{SXS}};
      \onslide<2>\node (img1 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.8cm,0.9cm)}]current page.center) {\includegraphics[height=6cm]{figs/Millisecond_NS2.png}}; 
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.8cm,3.2cm)}]current page.center) {Class.Quant.Grav. 30 (2013) 224003};
      \onslide<3>\node (img2 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.6cm,0.0cm)}]current page.center) {\includegraphics[height=4cm]{figs/NS_cross_section.png}}; 
      % Use Fig 3 of http://science.sciencemag.org/content/304/5670/536/tab-figures-data ?
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.6cm,-1.5cm)}]current page.center) {Robert Schulze};
      \onslide<4>\node (img2 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.6cm,0.0cm)}]current page.center) {\includegraphics[height=5cm]{figs/NS_temp.jpg}}; 
      %
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.6cm,2.2cm)}]current page.center) {Science 304, 5670, pp. 536-542};
       \onslide<5->\node (img3 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.cm,-2.8cm)}]current page.center){\includegraphics[height=3.8cm]{figs/HT_pulsar.jpg}};
       \onslide<6>\node (img3 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.2cm,0.9cm)}]current page.center){\includegraphics[height=5cm]{figs/HT_decay.jpg}}; 
       \node[anchor=center,scale=0.5,opacity=0.7] at ([shift={(2.5cm,-0.5cm)}]current page.center) {ApJ 829: 55 (2016)};
       %{\includegraphics[height=5cm]{figs/NS_EOS_Lattimer2012.png}}; 
      %\node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.cm,-3.8cm)}]current page.center) {Lattimer arXiv:1305.3510};
    \end{tikzpicture}
\end{frame}

\begin{frame}{Gravitational waves: Einstein 1916}
\vspace{-5mm}
\begin{center}
  \begin{tikzpicture}%[overlay,remember picture]
	      \onslide<1-4>\node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[height=3.5cm]{figs/Einstein1916.png}};
	      \onslide<2-4>{\draw[red,thick] (1.7,0.04) rectangle (3.1,0.27);}
  \end{tikzpicture}
  \\
  \vspace{5mm}
  \onslide<2->{Perturbations of spacetime with speed $=c$, sourced by accelerating masses.}
\end{center}
    %      
  \onslide<3->
  \[\text{Wave equation: } \boxed{\Box \bar{h}_{\mu\nu}= 0} \ \ \Longrightarrow
  \alert{\text{Plane-waves: }}  \bar{h}_{\mu\nu} = \Re\left[A_{\mu\nu}\, \text{e}^{-i\omega (t-z/\red{c} )}\right]
  \]
  %
 \onslide<4-> $A_{\mu\nu} $: polarization tensor, \\
 $\Longrightarrow$ \emph{only} \red{2} polarizations (\bl{transverse}):
  $\begin{array}{l} \w{A_{xx}=-A_{yy} = h_+}  \\ \w{A_{xy}=\ \ A_{yx} = h_\times} \\ \end{array}$
 \end{frame}

\begin{frame}{Gravitational waves: Einstein 1916}
\vspace{-5mm}
\begin{center}
  \begin{tikzpicture}%[overlay,remember picture]
	      \onslide<1>\node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[height=3.5cm]{figs/white_box.png}};
	      %\onslide<2-4>{\draw[red,thick] (1.7,0.04) rectangle (3.1,0.27);}
  \end{tikzpicture}
  \\
  \vspace{5mm}
  {Perturbations of spacetime with speed $=c$, sourced by accelerating masses.}
\end{center}
  %        
  %
  \[\text{Wave equation: } \boxed{\Box \bar{h}_{\mu\nu}= 0} \ \ \Longrightarrow
  \alert{\text{Plane-waves: }}  \bar{h}_{\mu\nu} = \Re\left[A_{\mu\nu}\, \text{e}^{-i\omega (t-z/\red{c} )}\right]
  \]
  %
 $A_{\mu\nu} $: polarization tensor, \\
 $\Longrightarrow$ \emph{only} \red{2} polarizations (\bl{transverse}):
  $\begin{array}{l}A_{xx}=-A_{yy} = \red{h_+}  \\ A_{xy}=\ \ A_{yx} = \red{h_\times}  \\ \end{array}$
 \begin{tikzpicture}[overlay,remember picture]
	      \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.5cm,1.5cm)}]current page.center){\animategraphics[loop,controls,height=2.6cm]{12}{GWs_plus-}{0}{20}};
	      \node (img1 )[anchor=center,scale=1.,opacity=1] at ([shift={(3.5cm,1.5cm)}]current page.center) {\animategraphics[loop,controls,height=2.6cm]{12}{GWs_cross-}{0}{20}};
\end{tikzpicture}
\end{frame}

%  Two of the same slide next for the right transition effects %%%%%%%%%%%%%%%%%%%%%%%

 \begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, GRBs, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave emission: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E}=L_\text{GW} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  \textcolor{white}{{\bf Binary systems:} 2 points masses in a circular orbit $M=m_1+m_2$}
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, GRBs, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave emission: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E}=L_\text{GW} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  {\bf Binary systems:} 2 points masses in a circular orbit $M=m_1+m_2$
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\animategraphics[loop,controls,height=2.9cm]{12}{Circ_Orb-}{0}{20}};
    %\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
    \onslide<2->\node (t1 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-1.3cm)}]current page.center) {$\mu=m_1 m_2/M^2,\ x(t) = r \cos\Omega t, \ y(t) = r \sin\Omega t $};
    \onslide<3->\node (t2 )[anchor=center,scale=1,opacity=1] at ([shift={(2.25cm,-2.2cm)}]current page.center) {$ Q^{ij} =  \mu\, x^i x^j, \qquad \boxed{\dot{E} = \f{32}{5}\f{G\mu^2}{c^5}r^4 \Omega^6}$};
    \onslide<4->\node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2.5cm,-3.2cm)}]current page.center) {Enhancement due to eccentricity \bl{$e$}};
    \node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-4cm)}]current page.center) {$ \bl{\f{\left(1 + \f{73}{24}e^2+ \f{37}{96} e^4 \right)}{(1-e^2)^{7/2}}}$};
  \end{tikzpicture}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Compact Binary Inspirals}
  \vspace{-4mm}
  Notation:
  \begin{itemize}
    \item $\red{f} $: GW frequency, \hspace{1.5cm} $\Longrightarrow\quad \omega\equiv2\pi f = 2\Omega$.
    %\item $\omega\equiv2\pi f = 2\Omega$: GW angular frequency.
    \item $\red{t}$: observer/detector time \quad$ \Longrightarrow\quad\dot{Y}= dY/dt$.
  \end{itemize}
  \onslide<2->
  {
    Assumptions:
    \begin{itemize}
      \item \bl{Quasi-circular} orbits: $e \ll 1$ and $\tfrac{|\dot{r}|}{r\Omega} <10^{-3} $ %at $f=100\,$Hz.
      \item \bl{Quadrupolar} GWs only: higher modes suppressed by $\tfrac{m_1-m_2}{M}\tfrac{v^2}{c^2}$
      \item Use Newtonian/post-Newtonian theory ($\le$3.5{\bf PN}).
      \vspace{0.5mm}
      \item \red{Terminate} inspiral at Schwarzschild \red{ISCO} 
      \begin{footnotesize}
      \begin{align*}
      f_\text{ISCO} &= \f{c^3}{\red{6}^{3/2}\pi G M} \simeq \bl{1571}\,\text{Hz}\, \left(\f{2.8M_\odot}{M}\right)\hspace{2cm}\\ 
      r_\text{ISCO} &\simeq 25\,\text{km}\left(\frac{M}{2.8 M_\odot}\right)
      \end{align*}
      \end{footnotesize}
    \end{itemize}
   } 
  %
  \onslide<3->
  {
    \vspace{1mm}
    For binary neutron stars ({\bf BNS})
    \begin{itemize}
      \item Point masses: tidal effects beyond $f \gtrsim 100\,$Hz
      \item Set $m_1 =m_2 = 1.4 M_\odot$, constraints: $1.2 M_\odot \lesssim m_\text{NS}\lesssim 1.6M_\odot$.
      \item Neglect spins: spin-orbit $\tfrac{\dot{E}_\text{SO}}{\dot{E}} \sim \tfrac{v_s}{c}\tfrac{v}{c}\tfrac{R_\text{NS}}{r}$, spin-spin $\sim \tfrac{v^4}{c^4}$
    \end{itemize}
  }
  \onslide<2->
  {
    \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.4cm,-1.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/OrbitPotential.pdf}};    
    \end{tikzpicture}
  }  
\end{frame}

\begin{frame}{Compact Binary Inspirals}{Newtonian evolution}
 Define \bl{chirp mass} \hspace{1.cm} $\red{M_c} \equiv \mu^{3/5} M^{2/5} = \tfrac{(m_1 m_2)^{3/5}}{(m_1+m_2)^{1/5}}$ \\
 %and symmetric mass ratio\qquad $\nu \equiv \tfrac{\mu}{M} =\tfrac{m_1 m_2}{M^2}$\\
 \vspace{1mm}
  which gives, via Kepler's \bl{third law}: $\Omega^2 = G M r^{-3}=\omega^2/4$,   
 \onslide<2->
 {
  \begin{footnotesize}
  $$ \boxed{\dot{E} = \f{32}{5}\f{c^5}{G} \left(\f{G M_c\, \omega}{2c^3}\right)^{10/3}}$$
  \end{footnotesize}\\
 } 
%
  \onslide<3->
  {
    {\bf Key idea:} \red{energy balance} between $\dot{E}$ and $\dot{E}_b$ (\bl{binding energy})\\
    %of gravitational binding energy
    \quad where $ E_b= -\tfrac{G m_1 m_2}{2r} = -\left(\tfrac{G^2 M_c^5 \omega^2}{32}\right)^{1/3} $
  }
%
  \onslide<4->
  {
    \[ 
    \dot{E}(f) = -\dot{E}_b(f) \Longrightarrow \boxed{\red{\dot{f}} = \f{96}{5}\pi^{8/3} \f{(G M_c)}{c^5}^{5/3}\, \red{f^{11/3}} }\hspace{2cm}
    \]
    \[
    \red{ \dot{f} \propto f^{11/3}\quad(1)}
    \]
    \begin{tikzpicture}[overlay,remember picture]
    \node[draw, red,thick,minimum width=3cm,minimum height=0.75cm] (b) at (5.9,0.85){};
    \end{tikzpicture}
  }

\end{frame}

\begin{frame}{Compact Binary Inspirals}
  \vspace{-3mm}
  Integrate $\dot{f} \sim f^{11/3} \quad \Longrightarrow \quad \int dt \sim \int f^{-11/3}df \quad\Rightarrow\quad t \sim \red{f^{-8/3}}$
  \\
  \vspace{1mm}
  Fix integration constant $\tau = t_\text{coal} - t >0$. \\
  \vspace{2mm}
  \onslide<2->
  Introduce \red{inspiral time}, i.e, time to \bl{coal}escence (merger)
  %
  \begin{small}
  \begin{align}
  \red{\tau_\text{insp}(f)} %&= \f{5}{256\pi}\f{c^5}{(\pi G M_c)^{5/3}} \,{f^{-8/3}}\nn\\
  &\simeq \bl{16.72\,\text{minutes}} \, \left(\f{1.219 M_\odot}{M_c}\right)^{5/3}\,{\left(\f{10\,\text{Hz}}{\red{f}}\right)^{\red{8/3}}\quad \red{(2)}}\nn\hspace{2.7cm}
  \end{align}
  \end{small}
  \\
  \onslide<3->
  {
    Number of GW cycles to coalescence: $\int \tfrac{f}{\dot{f}}df\sim \red{f^{-5/3}}$ %$\int df {f}/\dot{f}$
    \begin{small}
    \begin{align}
    \red{\mathcal{N}_\text{cyc}(f)} %&=\f{1}{32\pi} \f{c^5}{(\pi G M_c)^{5/3}}\left[ f^{-5/3}-f_\text{max}^{-5/3}\right]\nn \\
    &\approx 1.605\times \bl{10^4}\,\left(\f{1.219 M_\odot}{M_c}\right)^{5/3}{ \left(\f{10\,\text{Hz}}{\red{f}} \right)^{\red{5/3}}}\quad \red{(3)}\nn\hspace{3.2cm}
    \end{align}
    \end{small}
  }
  %
  \onslide<4->
      \begin{tikzpicture}[overlay,remember picture]
      \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.9cm,-3.25cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/R_of_Tau.pdf}};
      \onslide<5->\node[draw, red,thick,minimum width=9.5cm,minimum height=1.3cm] (b) at (4.4,3.35){};
      \onslide<5->\node[draw, red,thick,minimum width=9cm,minimum height=1.3cm] (b) at (4.1,1){};      
    \end{tikzpicture}\\
    %
    %\vspace{2mm}
    Binary separation \\
    \vspace{-1mm}
    \[ 
    \tfrac{\dot{r}}{r}\sim \tfrac{\dot{f}}{f} \Longrightarrow \red{r(\tau)}=r_i \left(\f{\red{\tau}}{\tau_i}\right)^{\red{1/4}} \hspace{4cm}
    \]
  %
\end{frame}

\begin{frame}{Gravitational waves from BNS}
  TT gauge, multipole, slow-velocity expansions give 
  \begin{small}
    \[
    h_{ij}^\text{TT} = \tfrac{2}{D}\tfrac{G}{c^4} \ddot{Q}_{ij}^\text{TT}\hspace{6cm}
    \]
    %
    \vspace{-5mm}
    \uncover<2->
    {
      \begin{align*}
      h_+(t) &= h_c(t)\, \left(\f{1+\cos^2\iota}{2}\right)\, \cos[\Phi(t)],\hspace{5cm}\\
      h_\times(t) & =h_c(t)\,\cos\iota \sin[\Phi(t)],
      \end{align*}
    }
  \end{small}
  %
  %
  \uncover<2->{$\iota =\,$ orbital \bl{inclination}}
  \vspace{-3mm}
  %
  \uncover<3->
  {
    \begin{small}
      \begin{align*}
      h_c(t) %&=\f{4}{D}\left(\f{G M_c}{c^2}\right)^{5/3}\, \left(\f{\pi f(t)}{c}\right)^{2/3}\nn\\
      &\approx 7.5\times \bl{10^{-24}} \left(\f{100\,\text{Mpc}}{D}\right)\left(\f{M_c}{1.219 M_\odot}\right)^{5/3}\left(\frac{\red{f}}{10\,\text{Hz}}\right)^{\red{2/3}} \\
      \Phi(t) &=\int_{t_i}^t dt' \omega(t') + \text{PN}
      \end{align*}
    \end{small}
    \\
  }
  %
  \uncover<4->
  {
    Frequency increases as $f^{11/3}$\\
    Amplitude increases as $f^{10/3} $
  }  
  %
  %
  \uncover<2->
  {
    \begin{tikzpicture}[overlay,remember picture]
      \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.5cm,1.2cm)}]current page.center) {\includegraphics[height=3cm]{figs/orbit_inc.png}};
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(3.7cm,-0.1cm)}]current page.center) {\textcolor{white}{\url{www.handprint.com/ASTRO/bineye3.html}}};
      \node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(2.7cm,-3.2cm)}]current page.center) {\includegraphics[height=2.5cm]{figs/white_box.png}};
      \node[anchor=center,scale=1.4,opacity=1] at ([shift={(1cm,-2.4cm)}]current page.center) {\href{https://www.youtube.com/watch?v=CH9JSqUfnSw}{\textcolor{white}{Chirp!}}};
    \end{tikzpicture}
  }
  \uncover<5->
  {
      \begin{tikzpicture}[overlay,remember picture]
      \node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(2.7cm,-3.2cm)}]current page.center) {\includegraphics[height=2.5cm]{figs/Chirp.pdf}};
      \node[anchor=center,scale=1.4,opacity=1] at ([shift={(1cm,-2.4cm)}]current page.center) {\href{https://www.youtube.com/watch?v=CH9JSqUfnSw}{\red{Chirp!}}};
    \end{tikzpicture}
  }  
\end{frame}

 \begin{frame}{GW Detectors: \red{Interferometers}}{Geodesic deviation}
    Two geodesics separated by $\xi^\mu$:
    %\[
    % \frac{D^2 \xi^\alpha}{D\tau^2} = -R^\alpha_{\beta \gamma\delta}\, \xi^\gamma u^\beta u^\delta 
    %\]
    $ \tfrac{d^2 \xi^i}{d\tau^2} = -R^i_{0j0} \xi^j\left(\tfrac{dx^0}{d\tau}\right)^2$ \hfill (detector frame)\\
    %\hspace{4cm} $\Longrightarrow \ddot{\xi^i} = -c^2 R^i_{0j0} \xi^j $ \hfill (detector time)\ \\
    \vspace{-3mm}
    \[
    \hspace{4.5cm} \boxed{\ddot{\xi^i} = \frac{1}{2} \ddot{h}_{ij}^\text{TT} \xi^j} \hspace{2.5cm} \text{(detector time)}
    \]
    %
    %
    \onslide<2->
    GW along $\mathbf{z}$: 
      \[
      h_{ij}^\text{TT} = \left(\begin{array}{cc}
				h_+ & h_\times \\
				h_\times & - h_+
				\end{array}\right) \cos\left(\omega\left( t-\tfrac{z}{c}\right)\right)
      \]
      \onslide<3->
      Consider mirrors at $(0,0)_1, (L,0)_2$ and $(L,L)_3$
      \begin{tikzpicture}[overlay,remember picture]
	%\node[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-3.4cm)}]current page.center){\todo{Make sure the vid here works}} ;
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.8cm,-2.8cm)}]current page.center) {\includegraphics[height=2.39cm]{figs/white_box.png}};
      \end{tikzpicture}
          \begin{small}
      \begin{align*}
	\w{s_{12} }& \w{=\int_1^2 \tfrac{ds}{dx}dx = (x_2-x_1) \left[ 1 + h_+ \cos(\omega t)\right]^{1/2}\hspace{4cm} }\\
	   &\w{ \approx L \left[ 1+\tfrac{1}{2} h_+ \cos\omega t \right] }\\
      \w{s_{13}} &\w{\approx L \left[ 1-\tfrac{1}{2} h_+ \cos\omega t \right]}
      \end{align*}
      \end{small}
    %
    %
 \end{frame}

 \begin{frame}{GW Detectors: \red{Interferometers}}{Geodesic deviation}
    Two geodesics separated by $\xi^\mu$:
    %\[
    % \frac{D^2 \xi^\alpha}{D\tau^2} = -R^\alpha_{\beta \gamma\delta}\, \xi^\gamma u^\beta u^\delta 
    %\]
    $ \tfrac{d^2 \xi^i}{d\tau^2} = -R^i_{0j0} \xi^j\left(\tfrac{dx^0}{d\tau}\right)^2$ \hfill (detector frame)\\
    %\hspace{4cm} $\Longrightarrow \ddot{\xi^i} = -c^2 R^i_{0j0} \xi^j $ \hfill (detector time)\ \\
    \vspace{-3mm}
    \[
    \hspace{4.5cm} \boxed{\ddot{\xi^i} = \frac{1}{2} \ddot{h}_{ij}^\text{TT} \xi^j} \hspace{2.5cm} \text{(detector time)}
    \]
    %
    %
    %
    GW along $\mathbf{z}$: 
      \[
      h_{ij}^\text{TT} = \left(\begin{array}{cc}
				h_+ & h_\times \\
				h_\times & - h_+
				\end{array}\right) \cos\left(\omega\left( t-\tfrac{z}{c}\right)\right)
      \]
      %
      Consider mirrors at $(0,0)_1, (L,0)_2$ and $(L,L)_3$
      \begin{tikzpicture}[overlay,remember picture]
	%\node[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-3.4cm)}]current page.center){\todo{Make sure the vid here works}} ;
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.8cm,-2.8cm)}]current page.center) {\animategraphics[loop,controls,height=2.39cm]{12}{LIGO_int-}{0}{102}};
      \end{tikzpicture}
    \onslide<2->      
      \begin{small}
      \begin{align*}
	s_{12} &= \int_1^2 \tfrac{ds}{dx}dx = (x_2-x_1) \left[ 1 + h_+ \cos(\omega t)\right]^{1/2}\hspace{4cm} \\
				    & \approx L \left[ 1+\tfrac{1}{2} h_+ \cos\omega t \right] \\
      s_{13} &\approx L \left[ 1-\tfrac{1}{2} h_+ \cos\omega t \right]
      \end{align*}
      \end{small}
    %
    %
 \end{frame}

\begin{frame}{Interferometers : \red{L}-topology}{LIGO, Virgo, KAGRA}
  \vspace{-2mm}
  Interferometers ({\bf IFO}s) measure only the response function (\bl{GW strain})
  \begin{footnotesize}
  \begin{align*}
   \bl{h(t)} &= F_+(\theta,\phi,\psi) \,h_+(t)+F_\times(\theta,\phi,\psi)\, h_\times(t) , \\
    F_+ &=\f{1}{2}\left(1+\cos^2\theta\right)\cos2\phi\cos 2\psi-\cos\theta \sin2\phi \sin2\psi, \\
    F_\times &=\f{1}{2}\left(1+\cos^2\theta\right)\cos2\phi\sin2\psi+\cos\theta \sin2\phi \cos2\psi
   \end{align*}
   \end{footnotesize}
   %
   \onslide<2->
   {
      Fourier transform to frequency domain 
      \begin{footnotesize}
      \begin{align*}
      &\red{\tilde{h}(f)} =A\, \underbrace{\f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6}}_{\bl{h_0}} \,\red{f^{-7/6}}\, e^{i\Psi}\, Q(\theta,\phi,\psi,\iota),\hspace{5cm} \\
	&\quad Q(\theta,\phi,\psi,\iota) = F_+\f{1+\cos^2\iota}{2}  + i F_\times \cos\iota \\
      &\quad A\equiv \pi^{-2/3}\sqrt{\tfrac{5}{24}}\\
      &\quad \Psi = \Psi_N(f) + \text{PN corrections}
	\end{align*}
	\end{footnotesize}
    }
    \onslide<3->
    {
      \red{Four} angles: $\{\theta,\phi,\iota,\psi \}$
      %
      \begin{tikzpicture}[overlay,remember picture]
      \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.8cm,-1.8cm)}]current page.center) {\includegraphics[height=5cm]{figs/Schutz_LRR_Fig3_half.jpg}};
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.1cm,-4.4cm)}]current page.center){Sathyaprakash, Schutz, Living Rev. Relativ. (2009) 12:2} ;
      \end{tikzpicture}
    }   
\end{frame}

\begin{frame}{Interferometers : \red{Triangular}-topology}{Einstein Telescope}
  \vspace{-2mm}
  \bl{Three} $60^\circ$ \red{V}-IFOs, one redundant (null stream), $L=10\,$km
  \onslide<2->
  {
    \begin{footnotesize}
    \begin{align*}
      &F^1_+(\theta,\phi,\psi)= \f{\sqrt{3}}{2} F_+(\theta,\phi,\psi), \quad F^1_\times(\theta,\phi,\psi)= \f{\sqrt{3}}{2} F_\times(\theta,\phi,\psi),\quad (\sin 60^\circ =\sqrt{3}/2)\\
      & F^2_{+,\times}(\theta,\phi,\psi) = F^1_{+,\times}(\theta,\phi+2\pi/3,\psi)\qquad (\text{rotation by 120}^\circ),\\
      & F^3_{+,\times}(\theta,\phi,\psi) = F^1_{+,\times}(\theta,\phi-2\pi/3,\psi) \qquad(\text{rotation by 240}^\circ),\\
      %&  |Q(\theta,\phi,\psi,\iota)|_\bigtriangleup^2 = \sum_{A=1}^3 \left(F^A_+\right)^2\f{1+\cos^2\iota}{2}  + \left(F^A_\times\right)^2 \cos\iota
      \end{align*}
      \end{footnotesize}
    }
   %
    \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.4cm,-1.cm)}]current page.center) {\includegraphics[height=3cm]{figs/gr_qc_12013563.png}};
    \node[anchor=center,scale=0.5,opacity=1] at ([shift={(5cm,0.4cm)}]current page.center) {Phys. Rev. D 86, 122001 };
    \end{tikzpicture}\\
    %
   \onslide<3->
   {
      No blind spots!
      \begin{footnotesize}
      \begin{align*}
      F^2_\bigtriangleup &\equiv \sum_{A=1}^3 \left(F^A_+\right)^2+\left(F^A_\times\right)^2 \hspace{8cm}\\ 
      &= \f{9}{32}\left(1+6\cos^2\theta + \cos^4\theta\right)
      \end{align*}
      \end{footnotesize}
  %
  %
  \begin{tikzpicture}[overlay,remember picture]

   \node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(0.4cm,-2.5cm)}]current page.center) {\includegraphics[height=4cm]{../Figures/ET_no_blind_spots.pdf}};
    \end{tikzpicture}
  }

\end{frame}

\begin{frame}{Detection criteria}{Optimal Signal-to-Noise Ratio ({\bf SNR})}
  \vspace{-5mm}
  \begin{footnotesize} 
  \[
  \rho=\left[ \int_0^\infty d\ln f\, \f{|\red{2\tilde{h}(f)\sqrt{f}}|^2}{S_n(f)}\right]^{1/2} 
  \]
  \\
  \end{footnotesize}
  $\sqrt{S_n(f)}\equiv$ {\it amplitude spectral density} ({\bf ASD}) or  detector noise
  \\
  \onslide<2->
  {
    \vspace{3mm}
    Recall, $\tilde{h}(f) \propto f^{-7/6}$ so $\tilde{h}(f) \sqrt{f} \propto \red{f^{-2/3}}$
    \begin{footnotesize} 
    \[
    \rho^2 =  (2A)^2 h_0^2 |Q(\theta,\phi,\psi,\iota)|^2 \int_0^{\bl{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)} 
    \]
    \end{footnotesize}
  }
  %
  \onslide<3->
  RMS \red{average} over the angles 
  \begin{small}
  $ \langle \ldots \rangle \equiv \int \sin\theta d\theta d\phi \int d\psi \int\sin\iota d\iota $
  \end{small}
  \begin{footnotesize} 
  \[
  \langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_\llcorner = \f{4}{25}, \qquad \langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_\triangle = \f{9}{25}
  \]
  \end{footnotesize}
  %
  Sources will have random $\{\theta,\phi,\psi,\iota\}$ \\
  \begin{footnotesize} 
  \[
  0 \le Q(\theta,\phi,\psi,\iota) \le \underbrace{Q_\text{max}(\theta,\phi,\psi,\iota)}_{\theta=\tfrac{\pi}{2},\psi=0,\iota=0}
  \]
  \end{footnotesize}
  % 
\end{frame}

\begin{frame}{L-IFO vs. $\bigtriangleup$-IFO Networks}
\red{Three L}-IFOs, e.g., LIGO L1, H1 + Virgo
%
  \begin{footnotesize} 
  \begin{align*}
  \rho_{\text{tot},\red{\llcorner}} &= \left[{\sum_{i=1}^3 \rho^2_i}\right]^{1/2} =\left[ \rho_h^2 + \rho_m^2 + \rho^2_l \right]^{1/2} 
	  = {\rho_c\left[\tilde{A}_h^2 + \tilde{A}_m^2 + \tilde{A}_l^2 \right]^{1/2}}\\
  \rho_c &\equiv 2 h_0 \left[ \int_0^{{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)}  \right]^{1/2}, \quad \left\{ \tilde{A}_h = \tfrac{4}{5}A,  \tilde{A}_m =\tfrac{2}{5}A,  \tilde{A}_l  < \tfrac{\tilde{A}_m}{2}\right\}
  \end{align*}
  \end{footnotesize}
  %
  \begin{tikzpicture}[overlay,remember picture]
  \node[draw, red,thick,minimum width=3.5cm,minimum height=0.8cm] (b) at (8.8,1.9){};
  \end{tikzpicture}
  \\
  %\vspace{2mm}
  \onslide<2->
  {
    $\bigtriangleup$ = \red{Three V-}IFOs: Einstein Telescope
    \begin{footnotesize} 
    \[
    \rho_{\text{tot},\bigtriangleup} = \left[\sum_{A=1}^3 \left(\rho^A\right)^2\right]^{1/2}  =  A \rho_c \,|Q(\theta,\phi,\psi,\iota)|_\bigtriangleup %\int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)} \,
    \]
    \end{footnotesize}
    \\
  }
  \onslide<3->
  Use \bl{RMS-averaged} $|Q|^2$
  \begin{footnotesize}
  \begin{align*}
  \rho_{\text{tot},\bigtriangleup} %&= \left[ \int_0^\infty d\ln f\, \f{|{2\tilde{H}_\bigtriangleup(f)\sqrt{f}}|^2}{S_n(f)}\right]^{1/2} 
  &=\f{6}{5}A\, h_0 \rho_c \\%\left[\int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)}\right]^{1/2} \\
  \end{align*}
  \end{footnotesize}
  \\
  %where \begin{small}$\tilde{H}_\bigtriangleup(f)\equiv A\,  \langle |Q|^2_\bigtriangleup\rangle\, h_0 f^{-7/6}= \f{3}{5} A h_0 f^{-7/6}$ \end{small}\\
  \vspace{-5mm}
  Compare with three-L network's \begin{small}$\rho_\text{tot} \approx \tfrac{4}{5}A\, h_0 \left[ \int df \ldots \right]^{1/2} $ \end{small}
  \vspace{10mm}
  %
  %
  \begin{tikzpicture}[overlay,remember picture]
  \node[draw, red,thick,minimum width=3.cm,minimum height=0.9cm] (b) at (-3.2,0.9){};
  \end{tikzpicture}
  %
  % 
\end{frame}

\begin{frame}{Case study: GW170817}
  \begin{itemize}
    \item 17 August 2017, signal swept from $f\approx \red{30}\,$Hz to $\bl{2000}\,$Hz 
    \onslide<2->\item The event lasted $\approx \red{57}\,$seconds, $\approx \bl{3000}$ GW cycles, $\rho_\text{tot} = \red{32.4}$
    \onslide<3->\item Followed by \red{GRB}170817A 1.7 seconds later
    \onslide<4->\item Rapid localization using L1-H1-V1 $\sim \bl{30}\,\text{deg}^2$
    \onslide<5->\item \red{Optical} counterpart in NGC 4993 11 hours later
  \end{itemize}

  \begin{columns}
  \begin{column}{0.5\textwidth}  %%<--- here
      \begin{center}
      \onslide<3->\includegraphics[height=4cm]{figs/GW_GRB170817.jpg}
      \end{center}
  \end{column}
  \begin{column}{0.5\textwidth}
  %
  \onslide<6->
  {
    Assuming low spin priors
    \begin{align*}
     M&=2.73^{+0.04}_{-0.01} M_\odot \\
     M_c &=1.186^{+0.001}_{-0.001}M_\odot  \\ 
     D&=40.7^{+2.36}_{-2.36}\text{Mpc}
    \end{align*}
  }
  \onslide<7->
  {
    \red{Note:} $M_c$ has the best precision: 
   \begin{small}$${\dot{f}} = \f{96}{5}\pi^{8/3} \left(\tfrac{G M_c}{c^3}\right)^{5/3}\, {f^{11/3}}$$ \end{small}
  }

    \end{column}
  \end{columns}

\end{frame}

\begin{frame}{Case study: GW170817}  {Let's put our knowledge to the \red{test}}
\vspace{-7mm}
  \begin{footnotesize}
    \begin{align*}
      &\tau_\text{insp}^{0\text{PN}}  \simeq 56.06\,\text{seconds}, \quad \mathcal{N}_\text{cyc}^{0\text{PN}} \simeq 2687 \hspace{2cm} (M=2.73 M_\odot) \\
      &\tau_\text{insp}^{2\text{PN}} \simeq 57.06\,\text{seconds}, \quad \mathcal{N}_\text{cyc}^{2\text{PN}} \simeq 2735 \hspace{2cm} (f_\text{ISCO}\simeq 1611\,\text{Hz}) 
    \end{align*}
  \end{footnotesize}
  \onslide<2->
  {
  What about the SNR? \quad
  \begin{small}$\rho_{\text{tot},\llcorner}  = {\bar\rho_c(f_0)\left[\tilde{A}_h^2 + \tilde{A}_m^2 + \tilde{A}_l^2 \right]^{1/2}}$   
  \end{small}
  }
  %

%
\begin{tikzpicture}[overlay,remember picture]
    \onslide<2->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-2cm,-2.cm)}]current page.center) {\includegraphics[height=5cm]{../Figures/GW170817_strains.pdf}};
    \onslide<2->\node (t1 )[anchor=center,scale=1,opacity=1] at ([shift={(3.65cm,0.8cm)}]current page.center) 
    {\begin{footnotesize} 
      $   \bar\rho_c(\red{f_0}) \equiv 2 h_0 \left[ \int_{\red{f_0}}^{{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)}  \right]^{1/2} $      
    \end{footnotesize}}; 
    \onslide<2->\node (t2 )[anchor=center,scale=1,opacity=1] at ([shift={(4.1cm,0.1cm)}]current page.center) 
    {\begin{footnotesize} 
      $ \tilde{A}_{i=h,m,l}=A\left\{\tfrac{4}{5},  \tfrac{2}{5},   < \tfrac{1}{5}\right\}$    
    \end{footnotesize}};  
    \onslide<3->\node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(4cm,-1.2cm)}]current page.center) 
    {\begin{footnotesize} 
      $\bar\rho_{i} = \left\{28.7, \ 14.3, \ 6.00 \right\}$    
    \end{footnotesize}}; 
    \onslide<4->\node (t4 )[anchor=center,scale=1,opacity=1] at ([shift={(4cm,-2.6cm)}]current page.center) 
    {\begin{footnotesize} 
      $\bar\rho_\text{LHV} = \{ 26.4, 18.8 , 2.0\} $    
    \end{footnotesize}};
    \onslide<3->\node (t5 )[anchor=center,scale=1,opacity=1] at ([shift={(4cm,-1.7cm)}]current page.center) 
    {\begin{footnotesize} 
      $\rho_\text{tot} = 32.0 $    
    \end{footnotesize}};
        \onslide<4->\node (t6 )[anchor=center,scale=1,opacity=1] at ([shift={(4cm,-3.1cm)}]current page.center) 
    {\begin{footnotesize} 
      $\rho_\text{tot} = 32.4 $    
    \end{footnotesize}};
    \onslide<3->\node[draw, red,thick,minimum width=3.8cm,minimum height=1.1cm] (b) at (10.1,-2.15){};
  \end{tikzpicture}
\end{frame}

\begin{frame}{GW Astronomy in the 2020s}{Advanced LIGO-Virgo}
  \red{BIG question:} how long before the GRB can we {\bf detect} BNS inspirals?\\
  Detection criterion: SNR = \red{15}. Set $m_1=m_2=1.4 M_\odot$\\
  Three identical L-IFOs with $Q_{i=h,m,l}$
  %\[  \bar\rho_\text{tot}= \sqrt{\left[\rho_h(f_0^h,\bar{f})\right]^2+\left[\rho_m(f_0^m,\bar{f})\right]^2+\left[\rho_l(f_0^l,\bar{f})\right]^2}=15\]
  % 
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<1->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.7cm)}]current page.center) {\includegraphics[height=5.5cm]{../Figures/ALigo_strains.pdf}};
      %\onslide<2->\node (t1 )[anchor=center,scale=1,opacity=1] at ([shift={(3.65cm,0.8cm)}]current page.center) 
      %{\begin{footnotesize} 
      %  $   \bar\rho_c(\red{f_0}) \equiv 2 h_0 \left[ \int_{\red{f_0}}^{{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)}  \right]^{1/2} $      
      %\end{footnotesize}}; 
      %\onslide<3->\node[draw, red,thick,minimum width=3.8cm,minimum height=1.1cm] (b) at (10.1,-2.15){};
  \end{tikzpicture}

\end{frame}

\begin{frame}{GW Astronomy in the 2020s}{Advanced LIGO-Virgo}
Detection moment: $t(f=\bar{f})$ such that
%There exists $\bar{f}$ such that
\begin{small}
  \[
    \bar\rho^2_\text{tot}= \left[\rho_h(f_0^h,\bar{f})\right]^2+\left[\rho_m(f_0^m,\bar{f})\right]^2+\left[\rho_l(f_0^l,\bar{f})\right]^2=\red{15}^2
  \]
\end{small}

  \begin{tikzpicture}[overlay,remember picture]
      \onslide<1->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.7cm)}]current page.center) {\includegraphics[height=5.5cm]{../Figures/ALigo_strains.pdf}};
      %\onslide<2->\node (t1 )[anchor=center,scale=1,opacity=1] at ([shift={(3.65cm,0.8cm)}]current page.center) 
      %{\begin{footnotesize} 
      %  $   \bar\rho_c(\red{f_0}) \equiv 2 h_0 \left[ \int_{\red{f_0}}^{{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)}  \right]^{1/2} $      
      %\end{footnotesize}}; 
      %\onslide<3->\node[draw, red,thick,minimum width=3.8cm,minimum height=1.1cm] (b) at (10.1,-2.15){};
  \end{tikzpicture}

\end{frame}

\begin{frame}{GW Astronomy in the 2020s}{Advanced LIGO-Virgo}
  Detection moment: $t(f=\bar{f})$ such that
  \begin{small}
    \[
	\bar\rho^2_\text{tot}= \left[\rho_h(f_0^h,\bar{f})\right]^2+\left[\rho_m(f_0^m,\bar{f})\right]^2+\left[\rho_l(f_0^l,\bar{f})\right]^2=\red{15}^2
    % \textcolor{black}{\bar\rho_\text{tot}=} \onslide<1>\left[\rho_h(f_0^h,\bar{f})\right]^2 \, \textcolor{white}{+}\onslide<2>\left[\rho_m(f_0^m,\bar{f})\right]^2 \, \textcolor{white}{+}\onslide<3->\left[\rho_l(f_0^l,\bar{f})\right]^2\textcolor{white}{=\red{15}^2}
    \]
  \end{small}

  \begin{tikzpicture}[overlay,remember picture]
      \onslide<1->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.7cm)}]current page.center) {\includegraphics[height=5.5cm]{../Figures/ALIGO2020_int_50Mpc.pdf}};
      \onslide<2->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.7cm)}]current page.center) {\includegraphics[height=5.5cm]{../Figures/ALIGO2020_int_100Mpc.pdf}};
      \onslide<3->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.7cm)}]current page.center) {\includegraphics[height=5.5cm]{../Figures/ALIGO2020_int_200Mpc.pdf}};
    \end{tikzpicture}

\end{frame}

\begin{frame}{GW Astronomy in the 2020s}{Advanced LIGO-Virgo}
  Advance \red{warning time}: inspiral time from $\bar{f}$
  \[
    T_\text{AW} = \tau_\text{insp}(\bar{f})
  \]
  \onslide<2->
  {
    Total accumulated SNR over the inspiral
    \begin{footnotesize}
      \[
      \bar{\rho}_\text{F}\equiv\sqrt{\left[{\rho}_h(f_0^h,{f}_\text{ISCO})\right]^2+\left[{\rho}_m(f_0^m,{f}_\text{ISCO})\right]^2+\left[{\rho}_l(f_0^l,{f}_\text{ISCO})\right]^2}
      \]
      \begin{table}[h]
	\centering
	\begin{tabular}{lcccccc}
	  \hline\hline
	  $D\,$(Mpc) & $\bar{f}\,$(Hz) &{}& $T^\text{0PN}_\text{AW}$(sec) &\hspace{1mm} & $T^\text{3.5PN}_\text{AW}\text{ (sec)}\hspace{2mm}$& $\bar{\rho}_F$\T \B \\
	  \hline
	  50 & $\approx\,$25.3 & & 84.25 & & 85.86 & 70.3 \\
	  100 & $\approx\,$41.6 & & 22.39 & & 22.86 & 35.2 \\
	  200 & $\approx\,$121 & & 1.291 &\quad & 1.300 & 17.6 \\
	  \hline\hline
	\end{tabular}
      \end{table}
    \end{footnotesize}
  }
%
\onslide<3->$\Longrightarrow$ NO likely FORECASTS  in the 2020s
\begin{tikzpicture}[overlay,remember picture]
  \onslide<1->\node[draw, red,thick,minimum width=2.8cm,minimum height=1.0cm] (b) at (-0.7,5.3){};
  \onslide<3->\node[draw, red,thick,minimum width=1.1cm,minimum height=1.2cm] (b) at (1.6,1.35){};
\end{tikzpicture}

\end{frame}

\begin{frame}{GW Astronomy in the 2030s}{The era of Einstein Telescope}
  \vspace{-1mm}
  \begin{itemize}
    \item[] One detector, \red{three V}-IFOs, $60^\circ$ 10\,km arms.
    \item[] Underground, cryogenic, \bl{xylophone} (high, low-frequency modes)
    \uncover<2->{
    \item[] Recall: \red{RMS}-average over all four angles
    \item[] Cosmological range $\Rightarrow$ \red{red}-shifted strain and SNR}
  \end{itemize}
  \onslide<2->
    \begin{footnotesize}
    \begin{align*} 
      \tilde{H}_\text{ET}(f) &= \f{3}{5}A h_0 \red{(1+z)^{-1}} f^{-7/6}\\
      \bar\rho_{\text{ET}}(f_0,f)&\equiv \f{6}{5}A h_0\, \red{(1+z)^{-1/6}}\left[\int_{f_0}^{f} d f'\, \f{f'^{-7/3}}{S^\text{ET}_n(f')}\right]^{1/2}
    \end{align*} 
    \end{footnotesize}
    \vspace{-1mm}
  \onslide<3->
    Consider BNS at
    \begin{footnotesize}
      $$ \bl{D}=\{100,200, 400, 600\}\,\text{Mpc} \Longrightarrow \red{z}=\{0.022, 0.044, 0.085, 0.12\}$$
      $$ \left(\text{Using } \Omega_\Lambda = 0.6911, \Omega_m = 0.3089, H_0 = 67.74\,\text{km}\,\text{s}^{-1}\,\text{Mpc}^{-1}\right)\qquad  $$
    \end{footnotesize}
  \onslide<4->
  {{\bf Note}: for binary black holes, ET's range $>\,1$Gpc, but inspirals too quick.}
\end{frame}

\begin{frame}{The era of Einstein Telescope}{BNS detections}
 Same detection criterion:\qquad $\bar{\rho}_\text{ET}(f_0, \bar{f}_\text{ET}) =15$\\
 \vspace{1mm}
 \red{Two} ET configurations:\qquad \textcolor{brown}{B} (high $f$), \red{C} (low $f$)
 \begin{tikzpicture}[overlay,remember picture]
      \onslide<1->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.5cm)}]current page.center) {\includegraphics[height=6cm]{../Figures/ET_strains_redshifted_pres.pdf}};
      \onslide<2->\node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.5cm)}]current page.center) {\includegraphics[height=6cm]{../Figures/ET_B_SNR_int.pdf}};
      \onslide<3->\node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.5cm)}]current page.center) {\includegraphics[height=6cm]{../Figures/ET_C_SNR_int.pdf}};
\end{tikzpicture}      
\end{frame}

\begin{frame}{The era of Einstein Telescope}{GRB \red{Forecast} Prospects } 
  \vspace{-8mm}
  \begin{footnotesize}
  \begin{table}[h]
  \centering
  \begin{tabular}{lccccccc}
  %toprule
  \hline\hline
  $D\,$(Mpc) & $R_\text{av}{}^\text{max}_\text{min} $ &  \multicolumn{3}{c}{ET-B}  & \multicolumn{3}{c}{ET-C}\T\B\\
  \hline
  {}& {} & $\bar{f}_\text{ET}\,$(Hz) & \hspace{2mm} $T_\text{AW}$ \hspace{2mm} & $\bar{\rho}_{F}$  & $\bar{f}_\text{ET}\,$(Hz) & \hspace{2mm} $T_\text{AW}\hspace{2mm}$& $\bar{\rho}_{F}$\T\B\\
  %midrule
  100 & $ 1.54^{4.7}_{0.3}$ & $\approx\,$6.72 &  47.0\,minutes & 306  & $\approx\,$3.27 & 5.34\,hours\ & 365\T\\
  200 &  $ 12.3^{38}_{2.6} $ & $\approx\,$11.2 & 11.6\,minutes & 152  & $\approx\,$4.10 & 2.87\,hours\ & 182 \\
  400 &$ 98.6^{300}_{20} $ & $\approx\,$18.2 & 3.00\,minutes & 75.7  & $\approx\,$5.06 & 1.51\,hours\ & 90.5\\
  600 & $ 333^{1020}_{69} $ & $\approx\,$26.0  &65.8\,seconds & 50.2 &    $\approx\,$5.7 & 1.03\,hours & 60.0  \\
  \hline\hline
  %bottomrule
  \end{tabular}
  \caption*{Event rates using $R=1540^{+3200}_{-1220}\,\text{Gpc}^{-3}\,\text{yr}^{-1}$ inferred from GW170817}
  \end{table}
  \end{footnotesize}
  %
  \vspace{-3mm}
  \onslide<2->Very \red{high SNRs} $\Rightarrow \ord(\rho_F^{\red{-1}})$-precision parameter estimation afterward \\
  \vspace{2mm}
  \onslide<3->\bl{ET-B} $\Longrightarrow $ forecasts not likely \\
  \vspace{2mm}
  \onslide<4->\red{ET-C} $\Longrightarrow T_\text{AW} \ge 1\,$hour within 600\,Mpc\\
  \vspace{1mm}
  \textcolor{white}{ET-C} $\Longrightarrow \red{\ord(100)}$ {\bf BNS forecasts} yearly! $\Rightarrow $ 1 per day!\\
  \vspace{1mm}
  \onslide<5->\textcolor{white}{ET-C} $\Longrightarrow \red{\ord(10)}$ with \red{3\,hours} of warning \\
  \vspace{1mm}
  \onslide<6->\textcolor{white}{ET-C} $\Longrightarrow \red{\ord(2)}$ with \red{5\,hours} of warning
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<3>\node[draw, blue,thick,minimum width=1.4cm,minimum height=1.6cm] (b) at (-2.6,5.3){};
      \onslide<4->\node[draw, red,thick,minimum width=1.8cm,minimum height=1.6cm] (b) at (3.6,5.3){};
  \end{tikzpicture}
\end{frame}

\begin{frame}{The era of Einstein Telescope}{Source localization before the merger} 
  ET is very sensitive, but all 3 IFOs are at one place. \\
  \vspace{1mm}
  $\Longrightarrow$ \red{poor} sky localization \\
  \vspace{1mm}
  \onslide<2->BUT \ldots \\
  \begin{itemize}
  \item With $T_\text{AW} \sim\,$hours, Earth's rotation matters!
  \item Two hours $\Rightarrow 30^\circ \Rightarrow \approx 2000\,$km of IFO distance $\sim\,$L1-H1 separation
  \end{itemize}
  \onslide<3->So, within 1\,Gpc, ET {\bf alone} will localize
  \begin{itemize}
  \item 20\% of the sources to within 100\,deg$^2$
  \item 2\% to within 20\,deg$^2$
  \item 0.5\% to within 10\,deg$^2$ 
  \onslide<4->\item[]$\Longrightarrow \red{5}$ to \red{10} sources within 400\,Mpc
  \end{itemize}
  \onslide<5->10\,deg$^2$: LSST's field of view\\
  \onslide<6-> Numbers improve with \bl{Cosmic Explorer} ($f\gtrsim \red{5}\,$Hz)
  %
  \begin{tikzpicture}[overlay,remember picture]
	\onslide<2->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.2cm,2.25cm)}]current page.center) {\includegraphics[height=3cm]{figs/Earth_rot.png}};
	\onslide<5->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.3cm,-2.25cm)}]current page.center) {\includegraphics[height=3.9cm]{figs/LSST.jpg}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Black hole - neutron star binaries}{Forecasting \red{tidal disruption} events}
  Necessary, but insufficient \bl{condition}:
  $   \f{M_\text{BH}}{m_\text{NS}} \gtrsim \left(\f{r}{R_\text{NS}}\right)^3  $
  %
  \onslide<2->
  \begin{itemize}
  \item Small separation $r\to r_\text{ISCO}$
  \item Low black hole mass, stellar evolution models: $M_\text{BH}\gtrsim 5M_\odot$
  \item High black hole spin: $\tilde{r}_\text{ISCO}\to 1$ from 6.
  \item Larger neutron star radii, $R_\text{NS}\sim15\,$km
  \end{itemize}
  \onslide<3->
  $\red{5} M_\odot -1.4 M_\odot$ BH-NS inspirals are \red{faster} because $\tau_\text{insp} \propto M^{-2/3}$
  %
  \onslide<4->
  \begin{footnotesize}
  \begin{table}[h]
  \centering
  \begin{tabular}{lccccc}
  %toprule
  \hline\hline
  $D\,$(Mpc) & \multicolumn{2}{c}{ET-B} &  & \multicolumn{2}{c}{ET-C}\T\B\\
  \hline
  {}&  \ \hspace{6mm}$T_\text{AW}$ \hspace{8mm} & $\bar{\rho}_{F}$ &{}  & $\hspace{6mm}T_\text{AW} \hspace{10mm}$ & $\bar{\rho}_{F}$\T\B\\
  %
  100 &   28.2\,minutes & 503 &{\qquad} &  2.15\,hours & 601\T\\
  200 & 8.21\,minutes & 251  &{\qquad} & 1.10\,hours & 300 \\
  %400 &  2.00\,minutes & 124 &{\qquad} & 34.2\,minutes & 148\\
  %1000 & 14.3\,seconds & 49 &{\qquad} & 13.9\,minutes& 58.5\\
  \hline\hline
  %bottomrule
  \end{tabular}
  \end{table}
  \end{footnotesize}
  %
  \onslide<5->
  $T_\text{AW} \gtrsim 1\,$hour $\Longrightarrow D \lesssim 200\,$Mpc \\
  O1 rate $ < 3600 \text{Gpc}^{-3}\text{yr}^{-1} \Longrightarrow 30\text{yr}^{-1}$ within $(200\,\text{Mpc})^3$ \\
  BH population studies: $\Longrightarrow \text{ \red{Forecast} } \red{3}\text{yr}^{-1}$ within $(200\,\text{Mpc})^3$
  % 
  %
  \begin{tikzpicture}[overlay,remember picture]
	  \onslide<1->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(5.6cm,2.25cm)}]current page.center) {\includegraphics[height=2.8cm]{figs/BH_NS_1.png}};
	  \node[anchor=center,scale=0.5,opacity=1] at ([shift={(5cm,0.7cm)}]current page.center) {Phys.Rev. D92 (2015) 044028};
  \end{tikzpicture}
\end{frame}

\begin{frame}{}{}%{Summary and Outlook}
{\bf Summary}
\begin{itemize}
 \item \bl{Quadrupole} + \bl{Newtonian} evolution = inspiral time $\tau_\text{insp}\propto \red{f^{-8/3}}$
 \item Detector sensitivity + GW strain = advance warning time $T_\text{AW}$
 \onslide<2->
 \item Advanced LIGO-Virgo will detect \bl{many} BNS, but NOT forecast.
 \onslide<3->
 \item Einstein Telescope will \bl{detect} $\bl{\ord(1000)}$ of BNS out to 1\,Gpc .
 \item Einstein-C will \red{forecast $\ord(100)$ BNS} with $T_\text{AW}>1\,$hour.
 \item Additionally forecast $< \ord(10)$ tidal disruption events.
 \item 5 to 10 BNS inspirals will be \red{localized} to $\red{10\,\text{deg}^2}$.
 \end{itemize}
 \onslide<4->{
  {\bf Outlook}
  \begin{itemize}
    \item Forecast/localize $\Rightarrow$ Mobilize EM telescopes \\
  \hspace{-5.6mm}$\Rightarrow$ \red{Observe} the merger/kilonova from  $\gamma$-rays to radio
  \item Pre-, post-merger image comparisons with LSST
 }
 \onslide<5->
 \item Contaminants? Type Ia supernovae, AGN, CV
\end{itemize}
\begin{tikzpicture}[overlay,remember picture]
	  \onslide<5->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-2.6cm,-3.8cm)}]current page.center) {\includegraphics[height=1.2cm]{figs/AAletter.png}};
	  \onslide<5->\node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(4.6cm,-3.cm)}]current page.center) {\includegraphics[height=2.7cm]{figs/SNIa_lim22mag.pdf}};
\end{tikzpicture}

\end{frame}

\begin{frame}{Supplement 1: mode suppression}
\begin{itemize}
 \item For $q=1$ due to symmetry only (2,2), (3,2), (4,2), (4,4) are nonzero 
 \item For $q\ne 1$ all $\ell \ge 2$ modes exist, but are much smaller than $(2,2)$
\end{itemize}
%\begin{align*}
%\end{align*}
\end{frame}

\begin{frame}{Supplement 2: The expression for $h_c(t)$}
Maggiore chapters 1, 3 and 4
\begin{itemize}
 \item Eq. (3.3) expanded to (3.9), (3.14) to (3.34) to (3.55) to (3.59)
 \item Eq. (3.72): $h(t)$ as angular distribution with mass moments $\ddot{M}_{ij}$'s
 \item At that point we get the scaling $h \sim \tfrac{1}{D}\tfrac{G}{c^4}\ddot{Q}_{ij}$.
 \item Eq. (3.275) shows it expanded in tensor harmonics so up to $\partial^2 Y^{\ell m}$
\end{itemize}
\end{frame}

\begin{frame}{Supp: coord. vs. proper distance in TT gauge}
  Geodesic deviation 
  \[
   \frac{D^2 \xi^\alpha}{D\tau^2} = -R^\alpha_{\beta \gamma\delta}\, \xi^\gamma u^\beta u^\delta 
  \]
In TT gauge,\qquad  $\Longrightarrow \quad\left.\tfrac{d^2 \xi^i}{d\tau^2}\right|_{\tau=0} = -\left. \dot{h}_{ij} \tfrac{d\xi^i}{d\tau}\right|_{\tau=0}$\\
i.e., if mirrors are initially at rest w.r.t to each other ( $\left. \tfrac{d\xi^i}{d\tau}\right|_{\tau=0} =0$ )\\
then $\tfrac{d^2 \xi^i}{d\tau^2}=0\quad \forall t$ \\
$\Longrightarrow$ coordinate separation $\xi^i = \text{constant}$ for all time \\
BUT, proper distance is NOT! \\
Consider mirrors at $\mathbf{x}=x_1$ and at $\mathbf{x}=x_2=x_1+\mathbf{\xi}$
\begin{align*}
  s = \int_1^2 \tfrac{ds}{dx}dx &= (x_2-x_1) \left[ 1 + h_+ \cos(\omega\t)\right]^{1/2} \\
			      & \approx L \left[ 1+\tfrac{1}{2} h_+ \cos\omega t \right] 
\end{align*}
 \end{frame}

\begin{frame}{Supplement 3: Direction of GW emission}
 Mostly off the orbital plane and slightly forward
 \begin{itemize}
  \item Eqs. (3.337-3.338) and Fig. 3.7 show that $\tfrac{d\dot{E}}{d\Omega}$ is max at the poles
  \item Eqs. (4.347), (4.348) with Fig. 4.24 (page 228) show that even for $v=0.99$ the beaming
  is much less significant compared to EM beaming shown in Fig. 4.23.
 \end{itemize}

\end{frame}

\end{document}

\begin{frame}{Supplemet 5: Mass-inclination degeneracy}

\end{frame}

\begin{frame}{Gravitational waves: d.o.f.'s}
$A_{\mu\nu} $ is the polarization tensor.
\begin{itemize}
 \uncover<2->{\item $\{ \bar{h}_{\mu\nu}, A_{\mu\nu} \}$: $4\times4$, symmetric \uncover<3->{$\hspace{1.5cm}\Rightarrow \f{4\times 5}{2}\ \ \ = 10$ d.o.f.}}
 \uncover<4->{\item Gauge condition: $\nabla_\mu \bar{h}^{\mu\nu} =  A_{\mu\nu} k^\nu= 0$ \uncover<5->{$\Rightarrow10-4 = \, 6$ d.o.f.}}
 \uncover<6->{\item Residual gauge freedom \hspace{2.25cm} \uncover<7->{$\Rightarrow 6-4 \ \ = \alert{2}$ \alert{d.o.f.}}}
 \end{itemize}
\uncover<7->{\quad \ Consistent with $\pm 2$ helicities of a \bl{massless} spin 2 boson.\\}
\uncover<8->{$\Rightarrow$ \alert{2} POLARIZATIONS (\alert{transverse}): \uncover<9->{$\begin{array}{l}A_{xx}=-A_{yy} = \red{h_+}  \\ A_{xy}=\ \ A_{yx} = \red{h_\times}  \\ \end{array}$}}
\\

\begin{tikzpicture}[overlay,remember picture]
  \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.5cm,-2.5cm)}]current page.center){\includegraphics[height=2.6cm]{figs/white_box.png}};
  \node (img1 )[anchor=center,scale=1.,opacity=1] at ([shift={(3.5cm,-2.5cm)}]current page.center) {\includegraphics[height=2.6cm]{figs/white_box.png}};
\end{tikzpicture}
\end{frame}

\begin{frame}{Gravitational waves: d.o.f.'s}
$A_{\mu\nu} $ is the polarization tensor.
\begin{itemize}
  {\item $\{ \bar{h}_{\mu\nu}, A_{\mu\nu} \}$: $4\times4$, symmetric {$\hspace{1.5cm}\Rightarrow \f{4\times 5}{2}\ \ \ = 10$ d.o.f.}}
  {\item Gauge condition: $\nabla_\mu \bar{h}^{\mu\nu} =  A_{\mu\nu} k^\nu= 0$ {$\Rightarrow10-4 = \, 6$ d.o.f.}}
  {\item Residual gauge freedom \hspace{2.25cm} {$\Rightarrow 6-4 \ \ = \alert{2}$ \alert{d.o.f.}}}
 \end{itemize}
{\quad \ Consistent with $\pm 2$ helicities of a \bl{massless} spin 2 boson.\\}
{$\Rightarrow$ \alert{2} POLARIZATIONS (\alert{transverse}): {$\begin{array}{l}A_{xx}=-A_{yy} = \red{h_+}  \\ A_{xy}=\ \ A_{yx} = \red{h_\times}  \\ \end{array}$}}
\\
%
\begin{tikzpicture}[overlay,remember picture]
  \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.5cm,-2.5cm)}]current page.center){\animategraphics[loop,controls,height=2.6cm]{12}{GWs_plus-}{0}{20}};
  \node (img1 )[anchor=center,scale=1.,opacity=1] at ([shift={(3.5cm,-2.5cm)}]current page.center) {\animategraphics[loop,controls,height=2.6cm]{12}{GWs_cross-}{0}{20}};
\end{tikzpicture}
\end{frame}